---
title: "p8105_hw1_xt2340"
author: "Xuening Tang"
date: "2026-09-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

```{r load_data}
library(tidyverse)
library(palmerpenguins)

```

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. Important variables in this dataset include `species` (e.g., Adelie, Chinstrap, Gentoo), `island` (Biscoe, Dream, Torgersen), `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`. The mean flipper length of the penguins in this dataset is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r penguin_scatter, warning=FALSE}
library(tidyverse)
library(palmerpenguins)
data("penguins", package = "palmerpenguins")

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Penguin Flipper Length vs Bill Length",
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)",
    color = "Species"
  )

ggsave("penguin_scatterplot.png", width = 6, height = 4)

```


# Problem 2

Create data frames

```{r prob2_dataframe}
library(tidyverse)

set.seed(8105)

prob2_df = tibble(
  samp = rnorm(10),
  samp_gt_0 = samp > 0,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  vec_factor = factor(c("Low", "Medium", "High", "Low", "Medium", "High", "Low", "Medium", "High", "Low"))
)

```

Take the mean of each column

```{r prob2_mean, eval=FALSE}
mean(pull(prob2_df, samp))
mean(pull(prob2_df, samp_gt_0))
mean(pull(prob2_df, vec_char))
mean(pull(prob2_df, vec_factor))

```


```{r prob2_as_numeric, eval=FALSE}
as.numeric(pull(prob2_df, samp_gt_0))
as.numeric(pull(prob2_df, vec_char))
as.numeric(pull(prob2_df, vec_factor))

```

Analysis:

Means can be successfully calculated for the numeric variable (`samp`) and the logical variable (`samp_gt_0`), where logical values are treated as 1 (`TRUE`) and 0 (`FALSE`).
Means cannot be calculated for the character variable (`vec_char`) and factor variable (`vec_factor`), resulting in `NA` values and warning messages.

